{\boldmath Pseudoscalar top-Higgs coupling: Exploration of $\mathrm{CP}$-odd observables to resolve the sign ambiguity }

### 2016-03-10 Paper_ttbarH_Draft.tex
KK: small edits to much of paper; also included a new diff.pdf file

%%%%%%%%%%%%%%%%% FIGURA 2 %%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{esquematico_II.pdf}
\vspace*{0.02\textwidth}
\caption{Schematic representation of the process $g_ag_b\to t(\to
  b_i{\ell}^+\nu_{\ell})\tbar(\to \bbar_j {\ell}^- \nubar) H$. The
  indices $i,j$ denote the colour of the quarks while $a,b$ are gluon
  indices.}
\label{fig2}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To derive the above expressions, we begin by considering the
schematic representation for the process $g_ag_b\to t(\to
b_i{\ell}^+\nu_{\ell})\,\tbar(\to \bbar_j {\ell}^- \nubar)\,H$
that is sketched in Fig.~\ref{fig2}.
Here $a$ and $b$ denote the initial-state gluons and
$i$ and $j$ refer to the colours of the top and antitop quarks.
%\par
%
The amplitude for this process may be written in the following
compact form
%
\beq
\label{eq6}
\mathcal{M}^{ab,ij}=\bar{\psi}_t\,\mathcal{A}^{ab,ij}\,\psi_{\tbar}\,,
\eeq
%
where the spinors $\bar{\psi}_t$ and $\psi_{\tbar}$
contain all of the information regarding the decay of the
virtual top and anti-top, respectively,
and where the quantity $A^{ab,ij}$ is given by
%
\beq
\label{eq5}
\mathcal{A}^{ab,ij}\equiv A^{ab,ij}_{\mu\nu}(\epsilon_{\lambda_a})^{\mu}(\epsilon_{\lambda_b})^{\nu}=\sum_{k=1}^8 \mathcal{A}^{ab,ij}_k =\kp \sum_{k=1}^8 \mathcal{S}^{ab,ij}_k + i\kpt \sum_{k=1}^8 \mathcal{P}^{ab,ij}_k.
\eeq 
%
The sum over $k$ in the above expression
corresponds to the eight gluon-initiated
diagrams indicated in Fig.~\ref{fig1}; also,
$\epsilon_{\lambda_a}$ and $\epsilon_{\lambda_b}$ are the
polarization vectors corresponding to $g_a$ and $g_b$, respectively.
In the last equality in Eq.~(\ref{eq5})
we have explicitly separated the amplitude into
two sums, with one sum corresponding to the scalar contributions
and the other to the pseudoscalar ones.
Taking all of the final-state particles to be massless, we can use the
spinor techniques
developed in \cite{Kleiss} to write $\bar{\psi}_t$ and
$\psi_{\tbar}$ as follows\footnote{These spinor techniques can also be used
  for massive final-state particles. Given the
  energy scale involved in the process in question, however, the assumption
  of massless final-state particles is sensible and greatly
  simplifies the derivation of Eq.~(\ref{eq2}).}
%
\bea
\label{eq7}
\bar{\psi}_t &=& -g^2\, \mathbb{P}_t(t)\,\mathbb{P}_W(t-b)\,
   \langle b-|\nu_{\ell}+\rangle \langle {\ell}^+\!+|(\tslash+m_t)\\
\label{eq8}
\psi_{\tbar} &=& g^2\, \mathbb{P}_t(\tbar)\,\mathbb{P}_W(\tbar-\bbar)\,
   \langle\nubar+|\bbar-\rangle (\tbslash-m_t)|{\ell}^-+\rangle,
\eea
%
where $|i+(-)\rangle \equiv (1/2)(1\pm \gamma^5)\,\psi_i$
represents a right-handed (left-handed) chiral spinor for final-state
particle $i$ and $\langle i+(-)|$ represents the corresponding adjoint spinor.
Also, $\mathbb{P}_t(q)=(q^2-m^2_t+im_t\Gamma_t)^{-1}$ and
$\mathbb{P}_W(q)=(q^2-m^2_W+im_W\Gamma_W)^{-1}$, and
we have denoted the momenta of the various particles by the symbols
that refer to the names of those particles~\cite{Mangano}.

Using the expressions defined above for $\bar{\psi}_t$ and
$\psi_{\tbar}$, we can write the amplitude $\mathcal{M}^{ab,ij}$
in a form that is (in a sense) factorized.
As a first step, we insert
Eqs.~(\ref{eq7}) and (\ref{eq8}) into Eq.~(\ref{eq6}), yielding
%
\beq
\label{eq9}
\mathcal{M}^{ab,ij}=\!-g^4\,\mathbb{P}_t(t)\,\mathbb{P}_t(\tbar)\,\mathbb{P}_W(t-b)\,\mathbb{P}_W(\tbar-\bbar)\,\langle b-|\nu_{\ell}+\rangle \langle\nubar+|\bbar-\rangle \sqrt{\strut2(t\cdot {\ell}^+)}\sqrt{\strut2(\tbar\cdot {\ell}^-)}\,[\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}],
%\frac{1}{(t^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{(\tbar^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{((t-b)^2-M^2_W)}\frac{1}{((\tbar-\bbar)^2-M^2_W)}\,\langle b-|\nu+\rangle\,\langle\nubar+|\bbar-\rangle
\eeq
%
where the spinors $\phi_{t}$ and $\phi_{\tbar}$ are defined as
%
\beq
\label{eq10}
\phi_t=\frac{(\tslash +m_t)}{\sqrt{\strut2(t\cdot {\ell}^+)}}|{\ell}^++\rangle
\eeq
%
\beq
\label{eq11}
\phi_{\tbar}=\frac{(\tbslash -m_t)}{\sqrt{\strut2(\tbar\cdot {\ell}^-)}}|{\ell}^-+\rangle \,.
\eeq
%
Note that in writing down the above expressions we have adopted the narrow-width
approximation for the top and antitop quarks and for the $W^\pm$ gauge
bosons.\footnote{Since Eq.~(\ref{eq9}) contains the top quark propagator
  term
  $\mathbb{P}_t(t)$, for example, $|\mathcal{M}^{ab,ij}|^2$ contains
  the factor $((t^2-m^2_t)^2+m^2_t\Gamma^2_t)^{-1}$,
  which is replaced by $(\pi/m_t\Gamma_t)\delta(t^2-m^2_t)$
  in the narrow-width approximation.  Thus, except for the propagator terms $\mathbb{P}_t(t)$ and $\mathbb{P}_t(\tbar)$,
   we take the four-vector $t$ appearing in Eqs.~(\ref{eq9})-(\ref{eq11}) to be
  on shell, satisfying $t^2 = m_t^2$.} 
Working out the projection operators $\phi_t\,\bar{\phi}_t$
and $\phi_{\tbar}\,\bar{\phi}_{\tbar}$, we have
%
\beq
\label{eq12}
\phi_t\,\bar{\phi}_t=\frac{1}{2}(1+\nslash_{\!t}\gamma^5)(\tslash +m_t)
\eeq
%
and
%
\beq
\label{eq13}
\phi_{\tbar}\,\bar{\phi}_{\tbar}=\frac{1}{2}(1+\nslash_{\!\tbar}\gamma^5)(\tbslash -m_t),
\eeq
%
with $n_t$ and $n_{\tbar}$ being the four-vectors defined in
Eqs.~(\ref{eq3}) and (\ref{eq4}).  Thus, $\phi_t$ and $\phi_{\tbar}$
may be regarded as
describing a top quark with spin vector $n_t$ and an antitop quark
with spin vector $n_{\tbar}$, respectively.

As a final step toward factorizing the amplitude $\mathcal{M}^{ab,ij}$,
we note that the amplitude for a top quark with spin vector
$n_t$ to decay into $b{\ell}^+\nu_{\ell}$ is given by
%
\beq
\label{eq14}
\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})=ig^2\mathbb{P}_W(t-b)\langle b-|\nu_{\ell}+\rangle\sqrt{\strut2(t\cdot {\ell}^+)} \,,
\eeq
%
and likewise,
%
\beq
\label{eq15}
\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})=ig^2\mathbb{P}_W(\tbar-\bar{b})\langle \bar{\nu}_{\ell}+|\bar{b}-\rangle\sqrt{\strut2(\tbar\cdot {\ell}^-)}.
\eeq
%
Furthermore, the term inside the square brackets in
Eq.~(\ref{eq9}) is the amplitude for producing a top quark
with spin vector $n_t$, along with an anti-top with spin vector
$n_{\tbar}$ and a Higgs boson,
%
\beq
\label{eq16}
\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H)=\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}.
\eeq
%
Combining Eqs.~(\ref{eq14})-(\ref{eq16}), we can write Eq.~(\ref{eq9})
in a form that appears to be factorized,
%
\beq
\label{eq17}
\mathcal{M}^{ab,ij}=\mathbb{P}_t(t)\mathbb{P}_t(\tbar)\,\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})\,\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})\,\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H) \,.
\eeq
%
It is important
to note that, even though the above expression has the appearance of being
factorized into production and decay parts, this apparent
factorization is a bit misleading.  In particular, the amplitude for
$t\tbar H$ production contains the top and antitop quark
spin four-vectors $n_t$ and $n_{\tbar}$, which
depend on final-state kinematical quantities [see
Eqs.~(\ref{eq3}) and (\ref{eq4})].
With this qualification in mind, we may now use the
amplitude in Eq.~(\ref{eq17})
to determine the corresponding scattering cross section.
After some manipulation of the phase space variables
to take advantage of the presence of the
propagator terms, $\mathbb{P}_t(t)$ and $\mathbb{P}_t(\tbar)$,
we arrive at the expression in Eq.~(\ref{eq2}).\footnote{The reader may note that in the differential widths of $t\to b{\ell}^+\nu_{\ell}$ and $\tbar\to \bar{b}{\ell}^-\bar{\nu}_{\ell}$ appearing in 
Eq.~(\ref{eq2}), the spin states of the top an antitop have been averaged. Interestingly, under the assumption of massless final-state particles, the amplitudes $\mathcal{M}(t(-n_t)\to b{\ell}^+\nu_{\ell})$ and $\mathcal{M}(\tbar(-n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})$ vanish. } This expression
also has the appearance of being factorized, but
qualifying remarks, similar to those above, apply.

@@

The expression derived above for the scattering cross
section [see Eq.~(\ref{eq2}), as well as Eq.~(\ref{eq17})]
provides significant insight into
how one might analyze $\ppprocess$ in order to determine
the nature of the top-Higgs coupling.  In particular,
let us focus on the production amplitude,
$\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H))$, which forms
part of the overall amplitude in Eq.~(\ref{eq17}).
The absolute value squared of the production amplitude
is used to determine $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$,
which in turn forms part of the expression for the ``factorized'' cross section
in Eq.~(\ref{eq2}).  Summing over colour and gluon indices
we have
%
\beq
\label{eq18}
\sum_{\substack{a,b \\ i,j}}|\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H)|^2=\sum_{\substack{a,b \\ i,j}}\left|\sum^{8}_{k=1}C^{ab,ij}_k\,\bar{\phi}_t (\kp\mathcal{S}_k+i\kpt\mathcal{P}_k)\phi_{\tbar}\right|^2,
\eeq
%
where we have separated the colour structure of each diagram by defining
$\mathcal{S}^{\,ab,ij}_k= C^{ab,ij}_k \mathcal{S}_k$ and
$\mathcal{P}^{\,ab,ij}_k= C^{ab,ij}_k \mathcal{P}_k$
[see Eqs.~(\ref{eq5}) and (\ref{eq16})]. Also, the factors
$g^2_s m_t/v$ and $-ig^2_s m_t/v$ arising from the vertices of the $t$-
and $s$-channel diagrams respectively have been included in the
definition of $C^{ab,ij}_k$ for convenience. The terms linear in $\kp$
and $\kpt$ can be written as
%
\beq
\label{eq19}
\mathcal{O}(\kp\kpt)\to \frac{1}{2}\kp\kpt \sum_{k,r}\mathbb{C}_{kr}\mathrm{Im}
\left\lbrace \mathrm{Tr}\left[ (1+\nslash_t \gamma^5)(\tslash+m_t)\mathcal{S}_k(1+\nslash_{\tbar}\gamma^5
 )(\tbslash -m_t)\tilde{\mathcal{P}}_r \right] \right\rbrace ,
\eeq
%
where the factor $\mathbb{C}_{kr}=\sum_{ab,ij}C^{ab,ij}_k
C^{ab,ij*}_r$ is real and where $\tilde{\mathcal{P}}_r = \gamma^0
\mathcal{P}^{\dagger}_r \gamma^0$.
The only terms that yield non-zero contributions
in the above sum are those with an
odd number of $\gamma^5$ matrices; these lead to triple-product
(TP) structures
of the form $\epsilon_{\alpha\beta\gamma\delta}\,p^{\alpha}_ap^{\beta}_bp^{\gamma}_cp^{\delta}_d$,
where $p_a$-$p_d$ represent various four momenta associated with the process.
In contrast,
it can be seen from Eq.~(\ref{eq18}) that the terms proportional to
$\kp^2$ and $\tilde{\kappa}^2_t$ descend from traces containing
an even number of $\gamma^5$
matrices and can be written in terms of scalar products of the
available momenta.\par
%\vspace*{-2mm}

With the above considerations in mind, it is useful to
write a general expression for the differential cross
section $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$
in terms of the momenta $q=(q_1-q_2)/2$,
$Q=(q_1+q_2)/2$, $t$, $\bar{t}$, $n_t$ and $n_{\tbar}$, where
$q_{1,2}$ denote the momenta of the initial-state gluons. Note that with this
choice, $q\cdot Q=0$ and $Q^2=-q^2=M^2_{t\tbar H}/4$, where $M_{t\tbar
  H}$ is the invariant mass of the $t\tbar H$ system.
Fifteen TPs can be constructed from these six
four-vectors,\footnote{We note that these fifteen
  TPs are not linearly independent (see the epsilon relations
  discussed in Ref.~\cite{identities}).} so that
%
\beq
\label{eq20}
d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)= \kp^2\,f_1(p_i\cdot p_j)+\tilde{\kappa}^2_t\,f_2(p_i\cdot p_j)+\kp\kpt\,\sum^{15}_{l=1}g_l(p_i\cdot p_j)\,\epsilon_l,
\eeq   
%
where
$\epsilon_l=\epsilon_{\alpha\beta\gamma\delta}\,p^{\alpha}_ap^{\beta}_bp^{\gamma}_cp^{\delta}_d$
denotes the $l$th TP (we adopt the convention $\epsilon_{0123}=+1$) and where $p_i$ and $p_j$ refer to any of the
six momenta.  The
functions $f_{1,2}$ and $g_k$ depend only on the possible scalar
products and are therefore even under a parity transformation
($\mathrm{P}$). However, the terms linear in $\kp\kpt$ are
$\mathrm{P}$-odd due to the presence of the $\mathrm{P}$-odd
TPs. Hence, only the functions $f_{1,2}$ will contribute to the total
cross-section, whereas the TP terms will be sensitive to the sign of
the anomalous coupling $\kpt$. Of the fifteen TPs mentioned above,
we will focus on those that contain both of the spin
vectors $n_t$ and $n_{\tbar}$, but do not include $q$.
The decision not to consider $q$-dependent TPs is motivated by the fact
that $q$ cannot be expressed in terms of the momenta of final state
particles (as $Q$ can, by virtue of energy-momentum conservation). The
decision to focus on TPs that contain both $n_t$ and $n_{\tbar}$
is rooted in the fact that the spins of pair-produced top and antitop quarks
are highly correlated at hadron colliders 
(even though the quarks themselves are unpolarized).
Observables that combine the decay products of the
$t$ and $\tbar$ will be sensitive to this spin
correlation~\cite{Bernreuther}.  A similar behaviour is expected in $t\tbar H$
production, where it can be shown that single-spin asymmetries
vanish~\cite{Ellis,Biswas}. Hence, in order to construct observables
sensitive to the structure of the $tH$ coupling, we will restrict our attention
to those
TPs that include information on the decay products of both the top and
anti-top quarks. Only five of the fifteen TPs in Eq.~(\ref{eq20}) do
not involve the four vector $q$ and, among these, only three
include both $n_t$ and $n_{\tbar}\,$.  Thus, we will restrict our attention
to the following TPs
%
\beq
\label{eqa1}
\epsilon_1\equiv\epsilon(t,\tbar,n_t,n_{\tbar}),\,\,
\vspace*{1mm}
\eeq
%
\beq
\label{eqa2}
\epsilon_2\equiv\epsilon(Q,\tbar,n_t,n_{\tbar}),
\vspace*{1mm}
\eeq
%
\beq
\label{eqa3}
\epsilon_3\equiv\epsilon(Q,t,n_t,n_{\tbar}).
\vspace*{1mm}
\eeq
%
\par

>>>>

%%%%%%%%%%%%%%%%% FIGURA 2 %%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{figure}[H]
\centering
\includegraphics[scale=0.6]{esquematico_II.pdf}
\vspace*{0.02\textwidth}
\caption{Schematic representation of the process $g_ag_b\to t(\to
  b_i{\ell}^+\nu_{\ell})\tbar(\to \bbar_j {\ell}^- \nubar) H$. The
  indices $i,j$ denote the colours of the quarks, while $a,b$ are gluon
  indices.}
\label{fig2}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To derive the above expressions, we begin by considering the
schematic representation for the process $g_ag_b\to t(\to
b_i{\ell}^+\nu_{\ell})\,\tbar(\to \bbar_j {\ell}^- \nubar)\,H$
that is sketched in Fig.~\ref{fig2}.
Here $a$ and $b$ denote the initial-state gluons and
$i$ and $j$ refer to the colours of the top and antitop quarks.
%\par
%
The amplitude for this process may be written in the following
compact form
%
\beq
\label{eq6}
\mathcal{M}^{ab,ij}=\bar{\psi}_t\,\mathcal{A}^{ab,ij}\,\psi_{\tbar}\,,
\eeq
%
where the spinors $\bar{\psi}_t$ and $\psi_{\tbar}$
contain all of the information regarding the decay of the
virtual top and anti-top, respectively,
and where the quantity $A^{ab,ij}$ is given by
%
\beq
\label{eq5}
\mathcal{A}^{ab,ij}\equiv A^{ab,ij}_{\mu\nu}(\epsilon_{\lambda_a})^{\mu}(\epsilon_{\lambda_b})^{\nu}=\sum_{k=1}^8 \mathcal{A}^{ab,ij}_k =\kp \sum_{k=1}^8 \mathcal{S}^{ab,ij}_k + i\kpt \sum_{k=1}^8 \mathcal{P}^{ab,ij}_k.
\eeq 
%
The sum over $k$ in the above expression
corresponds to the eight gluon-initiated
diagrams indicated in Fig.~\ref{fig1}; also,
$\epsilon_{\lambda_a}$ and $\epsilon_{\lambda_b}$ are the
polarization vectors corresponding to $g_a$ and $g_b$, respectively.
In the last equality in Eq.~(\ref{eq5})
we have explicitly separated the amplitude into
two sums, with one sum corresponding to the scalar contributions
and the other to the pseudoscalar ones.
Taking all of the final-state particles to be massless, we can use the
spinor techniques
developed in Ref.~\cite{Kleiss} to write $\bar{\psi}_t$ and
$\psi_{\tbar}$ as follows\footnote{These spinor techniques can also be used
  for massive final-state particles. Given the
  energy scale involved in the process in question, however, the assumption
  of massless final-state particles is sensible and greatly
  simplifies the derivation of Eq.~(\ref{eq2}).}
%
\bea
\label{eq7}
\bar{\psi}_t &=& -g^2\, \mathbb{P}_t(t)\,\mathbb{P}_W(t-b)\,
   \langle b-|\nu_{\ell}+\rangle \langle {\ell}^+\!+|(\tslash+m_t)\\
\label{eq8}
\psi_{\tbar} &=& g^2\, \mathbb{P}_t(\tbar)\,\mathbb{P}_W(\tbar-\bbar)\,
   \langle\nubar+|\bbar-\rangle (\tbslash-m_t)|{\ell}^-+\rangle,
\eea
%
where $|i+(-)\rangle \equiv (1/2)(1\pm \gamma^5)\,\psi_i$
represents a right-handed (left-handed) chiral spinor for final-state
particle $i$ and $\langle i+(-)|$ represents the corresponding adjoint spinor.
Also, $\mathbb{P}_t(q)=(q^2-m^2_t+im_t\Gamma_t)^{-1}$ and
$\mathbb{P}_W(q)=(q^2-m^2_W+im_W\Gamma_W)^{-1}$, and
we have denoted the momenta of the various particles by the symbols
that refer to the names of those particles~\cite{Mangano}.

Using the expressions defined above for $\bar{\psi}_t$ and
$\psi_{\tbar}$, we can write the amplitude $\mathcal{M}^{ab,ij}$
in a form that is (in a sense) factorized.
As a first step, we insert
Eqs.~(\ref{eq7}) and (\ref{eq8}) into Eq.~(\ref{eq6}), yielding
%
\beq
\label{eq9}
\mathcal{M}^{ab,ij}=\!-g^4\,\mathbb{P}_t(t)\,\mathbb{P}_t(\tbar)\,\mathbb{P}_W(t-b)\,\mathbb{P}_W(\tbar-\bbar)\,\langle b-|\nu_{\ell}+\rangle \langle\nubar+|\bbar-\rangle \sqrt{\strut2(t\cdot {\ell}^+)}\sqrt{\strut2(\tbar\cdot {\ell}^-)}\,[\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}],
%\frac{1}{(t^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{(\tbar^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{((t-b)^2-M^2_W)}\frac{1}{((\tbar-\bbar)^2-M^2_W)}\,\langle b-|\nu+\rangle\,\langle\nubar+|\bbar-\rangle
\eeq
%
where the spinors $\phi_{t}$ and $\phi_{\tbar}$ are defined as
%
\beq
\label{eq10}
\phi_t=\frac{(\tslash +m_t)}{\sqrt{\strut2(t\cdot {\ell}^+)}}|{\ell}^++\rangle
\eeq
%
\beq
\label{eq11}
\phi_{\tbar}=\frac{(\tbslash -m_t)}{\sqrt{\strut2(\tbar\cdot {\ell}^-)}}|{\ell}^-+\rangle \,.
\eeq
%
Note that in writing down the above expressions we have adopted the narrow-width
approximation for the top and antitop quarks and for the $W^\pm$ gauge
bosons{\bf (KK: remove ``\ldots and for the\ldots''?)}.\footnote{Since Eq.~(\ref{eq9}) contains the top quark propagator
  term
  $\mathbb{P}_t(t)$, for example, $|\mathcal{M}^{ab,ij}|^2$ contains
  the factor $((t^2-m^2_t)^2+m^2_t\Gamma^2_t)^{-1}$,
  which is replaced by $(\pi/m_t\Gamma_t)\delta(t^2-m^2_t)$
  in the narrow-width approximation.  Thus, except for the propagator terms $\mathbb{P}_t(t)$ and $\mathbb{P}_t(\tbar)$,
   we take the four-vector $t$ appearing in Eqs.~(\ref{eq9})-(\ref{eq11}) to be
  on shell, satisfying $t^2 = m_t^2$.} 
Working out the projection operators $\phi_t\,\bar{\phi}_t$
and $\phi_{\tbar}\,\bar{\phi}_{\tbar}$, we have
%
\beq
\label{eq12}
\phi_t\,\bar{\phi}_t=\frac{1}{2}(1+\nslash_{\!t}\gamma^5)(\tslash +m_t)
\eeq
%
and
%
\beq
\label{eq13}
\phi_{\tbar}\,\bar{\phi}_{\tbar}=\frac{1}{2}(1+\nslash_{\!\tbar}\gamma^5)(\tbslash -m_t),
\eeq
%
with $n_t$ and $n_{\tbar}$ being the four-vectors defined in
Eqs.~(\ref{eq3}) and (\ref{eq4}).  Thus, $\phi_t$ and $\phi_{\tbar}$
may be regarded as
describing a top quark with spin vector $n_t$ and an antitop quark
with spin vector $n_{\tbar}$, respectively.

As a final step toward factorizing the amplitude $\mathcal{M}^{ab,ij}$,
we note that the amplitude for a top quark with spin vector
$n_t$ to decay into $b{\ell}^+\nu_{\ell}$ is given by
%
\beq
\label{eq14}
\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})=ig^2\mathbb{P}_W(t-b)\langle b-|\nu_{\ell}+\rangle\sqrt{\strut2(t\cdot {\ell}^+)} \,,
\eeq
%
and likewise,
%
\beq
\label{eq15}
\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})=ig^2\mathbb{P}_W(\tbar-\bar{b})\langle \bar{\nu}_{\ell}+|\bar{b}-\rangle\sqrt{\strut2(\tbar\cdot {\ell}^-)}.
\eeq
%
Furthermore, the term inside the square brackets in
Eq.~(\ref{eq9}) is the amplitude for producing a top quark
with spin vector $n_t$, along with an anti-top with spin vector
$n_{\tbar}$ and a Higgs boson,
%
\beq
\label{eq16}
\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H)=\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}.
\eeq
%
Combining Eqs.~(\ref{eq14})-(\ref{eq16}), we can write Eq.~(\ref{eq9})
in a form that appears to be factorized,
%
\beq
\label{eq17}
\mathcal{M}^{ab,ij}=\mathbb{P}_t(t)\mathbb{P}_t(\tbar)\,\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})\,\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})\,\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H) \,.
\eeq
%
It is important
to note that, even though the above expression has the appearance of being
factorized into production and decay parts, this apparent
factorization is a bit misleading.  In particular, the amplitude for
$t\tbar H$ production contains the top and antitop quark
spin four-vectors $n_t$ and $n_{\tbar}$, which
depend on final-state kinematical quantities [see
Eqs.~(\ref{eq3}) and (\ref{eq4})].
With this qualification in mind, we may now use the
amplitude in Eq.~(\ref{eq17})
to determine the corresponding scattering cross section.
After some manipulation of the phase space variables
to take advantage of the presence of the
propagator terms, $\mathbb{P}_t(t)$ and $\mathbb{P}_t(\tbar)$,
we arrive at the expression in Eq.~(\ref{eq2}).\footnote{The reader may note that in the differential widths of $t\to b{\ell}^+\nu_{\ell}$ and $\tbar\to \bar{b}{\ell}^-\bar{\nu}_{\ell}$ appearing in 
Eq.~(\ref{eq2}), the spin states of the top and antitop have been averaged. Interestingly, under the assumption of massless final-state particles, the amplitudes $\mathcal{M}(t(-n_t)\to b{\ell}^+\nu_{\ell})$ and $\mathcal{M}(\tbar(-n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})$ vanish. } This expression
also has the appearance of being factorized, but
qualifying remarks, similar to those above, apply.

@@

The expression derived above for the scattering cross
section [see Eq.~(\ref{eq2}), as well as Eq.~(\ref{eq17})]
provides significant insight into
how one might analyze $\ppprocess$ in order to determine
the nature of the top-Higgs coupling.  In particular,
let us focus on the production amplitude,
$\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H))$, which forms
part of the overall amplitude in Eq.~(\ref{eq17}).
The absolute value squared of the production amplitude
is used to determine $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$,
which in turn forms part of the expression for the ``factorized'' cross section
in Eq.~(\ref{eq2}).  Summing over colour and gluon indices
we have
%
\beq
\label{eq18}
\sum_{\substack{a,b \\ i,j}}|\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H)|^2=\sum_{\substack{a,b \\ i,j}}\left|\sum^{8}_{k=1}C^{ab,ij}_k\,\bar{\phi}_t (\kp\mathcal{S}_k+i\kpt\mathcal{P}_k)\phi_{\tbar}\right|^2,
\eeq
%
where we have separated the colour structure of each diagram by defining
$\mathcal{S}^{\,ab,ij}_k= C^{ab,ij}_k \mathcal{S}_k$ and
$\mathcal{P}^{\,ab,ij}_k= C^{ab,ij}_k \mathcal{P}_k$
[see Eqs.~(\ref{eq5}) and (\ref{eq16})]. Also, the factors
$g^2_s m_t/v$ and $-ig^2_s m_t/v$ arising from the vertices of the $t$-
and $s$-channel diagrams, respectively, have been included in the
definition of $C^{ab,ij}_k$ for convenience. The terms linear in $\kp$
and $\kpt$ can be written as
%
\beq
\label{eq19}
\mathcal{O}(\kp\kpt)\to \frac{1}{2}\kp\kpt \sum_{k,r}\mathbb{C}_{kr}\mathrm{Im}
\left\lbrace \mathrm{Tr}\left[ (1+\nslash_t \gamma^5)(\tslash+m_t)\mathcal{S}_k(1+\nslash_{\tbar}\gamma^5
 )(\tbslash -m_t)\tilde{\mathcal{P}}_r \right] \right\rbrace ,
\eeq
%
where the factor $\mathbb{C}_{kr}=\sum_{ab,ij}C^{ab,ij}_k
C^{ab,ij*}_r$ is real and where $\tilde{\mathcal{P}}_r = \gamma^0
\mathcal{P}^{\dagger}_r \gamma^0$.
The only terms that yield non-zero contributions
in the above sum are those with an
odd number of $\gamma^5$ matrices; these lead to triple-product
(TP) structures
of the form $\epsilon_{\alpha\beta\gamma\delta}\,p^{\alpha}_ap^{\beta}_bp^{\gamma}_cp^{\delta}_d$,
where $p_a$-$p_d$ represent various four momenta associated with the process.
In contrast,
it can be seen from Eq.~(\ref{eq18}) that the terms proportional to
$\kp^2$ and $\tilde{\kappa}^2_t$ descend from traces containing
an even number of $\gamma^5$
matrices and can be written in terms of scalar products of the
available momenta.\par
%\vspace*{-2mm}

With the above considerations in mind, it is useful to
write a general expression for the differential cross
section $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$
in terms of the momenta $q=(q_1-q_2)/2$,
$Q=(q_1+q_2)/2$, $t$, $\bar{t}$, $n_t$ and $n_{\tbar}$, where
$q_{1,2}$ denote the momenta of the initial-state gluons. Note that with this
choice, $q\cdot Q=0$ and $Q^2=-q^2=M^2_{t\tbar H}/4$, where $M_{t\tbar
  H}$ is the invariant mass of the $t\tbar H$ system.
Fifteen TPs can be constructed from these six
four-vectors,\footnote{We note that these fifteen
  TPs are not linearly independent (see the epsilon relations
  discussed in Ref.~\cite{identities}).} so that
%
\beq
\label{eq20}
d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)= \kp^2\,f_1(p_i\cdot p_j)+\tilde{\kappa}^2_t\,f_2(p_i\cdot p_j)+\kp\kpt\,\sum^{15}_{l=1}g_l(p_i\cdot p_j)\,\epsilon_l,
\eeq   
%
where
$\epsilon_l=\epsilon_{\alpha\beta\gamma\delta}\,p^{\alpha}_ap^{\beta}_bp^{\gamma}_cp^{\delta}_d$
denotes the $l$th TP (we adopt the convention $\epsilon_{0123}=+1$) and where $p_i$ and $p_j$ refer to any of the
six momenta.  The
functions $f_{1,2}$ and $g_k$ depend only on the possible scalar
products and are therefore even under a parity transformation
($\mathrm{P}$). However, the terms linear in $\kp\kpt$ are
$\mathrm{P}$-odd due to the presence of the $\mathrm{P}$-odd
TPs. Hence, only the functions $f_{1,2}$ will contribute to the total
cross section, whereas the TP terms will be sensitive to the sign of
the anomalous coupling $\kpt$. Of the fifteen TPs mentioned above,
we will focus on those that contain both of the spin
vectors $n_t$ and $n_{\tbar}$, but do not include $q$.
The decision not to consider $q$-dependent TPs is motivated by the fact
that $q$ cannot be expressed in terms of the momenta of final state
particles (as $Q$ can, by virtue of energy-momentum conservation). The
decision to focus on TPs that contain both $n_t$ and $n_{\tbar}$
is rooted in the fact that the spins of pair-produced top and antitop quarks
are highly correlated at hadron colliders 
(even though the quarks themselves are unpolarized).
Observables that combine the decay products of the
$t$ and $\tbar$ will be sensitive to this spin
correlation~\cite{Bernreuther}.  A similar behaviour is expected in $t\tbar H$
production, where it can be shown that single-spin asymmetries
vanish~\cite{Ellis,Biswas}. Hence, in order to construct observables
sensitive to the structure of the $tH$ coupling, we will restrict our attention
to those
TPs that include information on the decay products of both the top and
anti-top quarks. Only five of the fifteen TPs in Eq.~(\ref{eq20}) do
not involve the four vector $q$ and, among these, only three
include both $n_t$ and $n_{\tbar}\,$.  Thus, we will restrict our attention
to the following TPs
%
\beq
\label{eqa1}
\epsilon_1\equiv\epsilon(t,\tbar,n_t,n_{\tbar}),\,\,
\vspace*{1mm}
\eeq
%
\beq
\label{eqa2}
\epsilon_2\equiv\epsilon(Q,\tbar,n_t,n_{\tbar}),
\vspace*{1mm}
\eeq
%
\beq
\label{eqa3}
\epsilon_3\equiv\epsilon(Q,t,n_t,n_{\tbar}).
\vspace*{1mm}
\eeq
%
\par
